---
title: "Part 2 - Chi Squared w/ Lizards Tails"
author: "Max Settineri"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(broom)

```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```
Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```
## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grasslands), RABB (rabbit mesquite site), EAST (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_count <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_count %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
# default is to calculate % by rows
  
```

## Covert to a contingency table

use `column_to_rowname()`.

```{r}
lizard_ct <- tail_count %>% 
  column_to_rownames(var = "site")
```

## Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```
Using a significance level of .05, yes we have sufficient evidence to reject null hypothesis so we retain alternative hypothesis that tail size and breakage are not independent. 

There is a significant association (ie. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
